---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Overview

<!-- badges: start -->
<!-- badges: end -->

The goal of text2speech is to harmonize various text-to-speech engines, including Amazon Polly, Coqui TTS, Google Cloud Text-to-Speech API, and Microsoft Cognitive Services Text to Speech REST API. 

All of these engines (except Coqui TTS) are made available as R packages:

- [aws.polly](https://github.com/cloudyr/aws.polly) is a client for [Amazon Polly](https://docs.aws.amazon.com/polly/latest/dg/what-is.html)
- [googleLanguageR](https://github.com/ropensci/googleLanguageR) is a client to the [Google Cloud Text-to-Speech API](https://cloud.google.com/text-to-speech/) 
- [mscstts2](https://github.com/howardbaek/mscstts2) is a client to the [Microsoft Cognitive Services Text to Speech REST API](https://learn.microsoft.com/en-us/azure/cognitive-services/speech-service/rest-text-to-speech?tabs=streaming)

Coqui TTS functionality is included directly in this package.


## Installation

You can install this package from CRAN or the development version from GitHub with:

``` r
# Install from CRAN
install.packages("text2speech")

# or the development version from GitHub
# install.packages("devtools")
devtools::install_github("jhudsl/text2speech")
```

## Authentication

Check for authentication (permission to use each of the different text-to-speech engines). If not authenticated, this needs to be set up in each service separately by the user.

```{r example}
library(text2speech)

# Amazon Polly
tts_auth("amazon")
# Coqui TTS
tts_auth("coqui")
# Google Cloud Text-to-Speech API 
tts_auth("google")
# Microsoft Cognitive Services Text to Speech REST API
tts_auth("microsoft")
```


## Voices

List out different voice options for each service

```{r}
# Amazon Polly
voices_amazon <- tts_amazon_voices()
head(voices_amazon)

# Coqui TTS
voices_coqui <- tts_coqui_voices()
head(voices_coqui)

# Google Cloud Text-to-Speech API 
voices_google <- tts_google_voices()
head(voices_google)

# Microsoft Cognitive Services Text to Speech REST API
voices_microsoft <- tts_microsoft_voices()
head(voices_microsoft)
```

## Convert text to speech

```r
# Amazon Polly
tts("Hello world!", service = "amazon")

# Coqui TTS
tts("Hello world!", service = "coqui")

# Google Cloud Text-to-Speech API 
tts("Hello world!", service = "google")

# Microsoft Cognitive Services Text to Speech REST API
tts("Hello world!", service = "microsoft")
```

