---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Overview

<!-- badges: start -->
<!-- badges: end -->

The goal of text2speech is to harmonize various text-to-speech engines, including Amazon Polly, Coqui TTS, Google Cloud Text-to-Speech API, and Microsoft Cognitive Services Text to Speech REST API. 

With the exception of Coqui TTS, all these engines are accessible as R packages:

- [aws.polly](https://github.com/cloudyr/aws.polly) is a client for [Amazon Polly](https://docs.aws.amazon.com/polly/latest/dg/what-is.html)
- [googleLanguageR](https://github.com/ropensci/googleLanguageR) is a client to the [Google Cloud Text-to-Speech API](https://cloud.google.com/text-to-speech/) 
- [conrad](https://github.com/fhdsl/conrad) is a client to the [Microsoft Cognitive Services Text to Speech REST API](https://learn.microsoft.com/en-us/azure/cognitive-services/speech-service/rest-text-to-speech?tabs=streaming)


You might notice Coqui TTS doesn't have its own R package. This is because, at this time, text2speech directly incorporates the functionality of Coqui TTS. The R wrapper of Coqui is [under development](https://github.com/howardbaek/ribbit).


## Installation

You can install this package from CRAN or the development version from GitHub with:

``` r
# Install from CRAN
install.packages("text2speech")

# or the development version from GitHub
# install.packages("devtools")
devtools::install_github("jhudsl/text2speech")
```

## Authentication

Check for authentication. If not already authenticated, users must individually configure it for each service.

```{r example}
library(text2speech)

# Amazon Polly
tts_auth("amazon")
# Coqui TTS
tts_auth("coqui")
# Google Cloud Text-to-Speech API 
tts_auth("google")
# Microsoft Cognitive Services Text to Speech REST API
tts_auth("microsoft")
```


## Voices

List different voice options for each service.

```{r}
# Amazon Polly
voices_amazon <- tts_amazon_voices()
head(voices_amazon)

# Coqui TTS
voices_coqui <- tts_coqui_voices()
head(voices_coqui)

# Google Cloud Text-to-Speech API 
voices_google <- tts_google_voices()
head(voices_google)

# Microsoft Cognitive Services Text to Speech REST API
voices_microsoft <- tts_microsoft_voices()
head(voices_microsoft)
```

## Convert text to speech

Synthesize speech with `tts(text = "TEXT", service = "ENGINE")`

```r
# Amazon Polly
tts("Hello world!", service = "amazon")

# Coqui TTS
tts("Hello world!", service = "coqui")

# Google Cloud Text-to-Speech API 
tts("Hello world!", service = "google")

# Microsoft Cognitive Services Text to Speech REST API
tts("Hello world!", service = "microsoft")
```

The resulting output will consist of a standardized tibble featuring the following columns:

- `index`: Sequential identifier number
- `original_text`: The text input provided by the user
- `text`: In case `original_text` exceeds the character limit, `text` represents the outcome of splitting `original_text`. Otherwise, `text` remains the same as `original_text`.
- `wav`: Wave object (S4 class)
- `file`: File path to the audio file
- `audio_type`: The audio format, either mp3 or wav
- `duration` : The duration of the audio file
- `service`: The text-to-speech engine used

