---
title: "Coqui TTS"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Coqui TTS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(text2speech)
```

## Introduction

Coqui TTS is a text-to-speech (TTS) library that enables the conversion of regular text into speech. It provides pre-trained models and vocoders as part of its package.

The underlying technology of text-to-speech is highly intricate. However, if you're interested in delving deeper into the subject, here are some recommended talks:

- [Pushing the frontier of neural text to speech](https://youtu.be/MA8PCvmr8B0), a webinar by Xu Tan at Microsoft Research Asia
- [Text to Speech Deep Dive](https://www.youtube.com/watch?v=aLBedWj-5CQ), a talk given during the ML for Audio Study Group hosted by Hugging Face.




## Authentication

```{r}

```


## List Voices

```{r}

```


## Text-to-Speech

```{r}

```





